---
title: "R beyond Data Analysis"
subtitel: "Function Plotting"
author: 
- name: "Charlotte Haussmann"
  url: https://github.com/LotteHaussmann/R-beyond-data-analysis
  affiliation: Uni.lu
  affiliation_url: https://www.jugend-in-luxemburg.lu
date: "`r Sys.Date()`"
output: 
  html_document: default
abstract: Here we learn how to write a *plotting function*.
keywords: ["R", "Function", "rmarkdown", "yaml"]
subject: Social Science
lang: "en"
---

```{r setup, include=FALSE}
```

```{r, echo=F, warning=FALSE, message=FALSE}
r <- getOption("repos")
r["CRAN"] <-"https://cloud.r-project.org/"
options(repos=r)

if (!require(haven)) {
    install.packages("haven")
    require(haven)
}

if (!require(ggplot2)) {
    install.packages("ggplot2")
    require(ggplot2)
}
if (!require(tidyverse)) {
    install.packages("tidyverse")
    require(tidyverse)
}
```

  
***  
### How to write a plotting function 
  
First we read load the packages *haven* and *ggplot2*. 
```{r, echo=T,eval=F}
library(haven)
library(ggplot2)

```

The next step is to load the data from Stanciu et al. (2017). 
```{r, echo = T}
df <- read_sav("data/sample.sav")
```


After that we are creating a function for plotting. 
```{r, echo = FALSE}
func_plot <- function(data, variable){
  
  if (!(variable %in% names(data))) {
    stop("Variable not found in data frame")
  }
 
  if(!is.numeric(data[[variable]])) {
    stop("Variable is not numeric")
  }
  
  ggplot(data, aes_string(x = variable)) +
    geom_bar(fill = "skyblue", color = "black") +
    labs(title = paste("Histogram of", variable),
         x = variable,
         y = "Frequency") +
    theme_minimal()
  
}
```
***
##### Let's look at it, step by step.  
We create a new object **func_plot**. This function accesses a **data set** and the **variables** contained in it  
```{r, echo = T, eval = F}
func_plot <- function(data, variable){
  
}
```

First we want to check if the called variable is inside the data set. If not, there will be a warning message displayed: *"Variable not found in data frame"*.

```{r, echo = T, eval = F}
func_plot <- function(data, variable){
  
  if (!(variable %in% names(data))) {
    stop("Variable not found in data frame")
  }}
```
After that we check if the variable is or is not numeric. If the variable is **not** numeric, the warning message *"Variable is not numeric"* will be displayed. 
```{r, echo = T, eval = F}
func_plot <- function(data, variable){
  
  if (!(variable %in% names(data))) {
    stop("Variable not found in data frame")
  }
 
  if(!is.numeric(data[[variable]])) {
    stop("Variable is not numeric")
  }}
```
The last step is to great a plot with the package *ggplot2*. We are creating a histogram with *geom_bar* and customized colours and title.  
```{r}
func_plot <- function(data, variable){
  
  if (!(variable %in% names(data))) {
    stop("Variable not found in data frame")
  }
 
  if(!is.numeric(data[[variable]])) {
    stop("Variable is not numeric")
  }
  
  ggplot(data, aes_string(x = variable)) +
    geom_bar(fill = "skyblue", color = "black") +
    labs(title = paste("Histogram of", variable),
         x = variable,
         y = "Frequency") +
    theme_minimal()
  
}
```

Now we will use this new function with the variable **gen** which stands for gender, with ($1=female;2=male$). 
```{r, warning=FALSE, message=FALSE, out.width = '60%'}
func_plot(df, "gen")
```

  
And once again, with the variable **age**. 
```{r, out.width = '120%'}
func_plot(df, "age")
```

  
And a last time with **men_warm**.
```{r, warning=FALSE, message=FALSE}
func_plot(df, "men_warm") 
```

***
***
## Plain vs. live text 
```{r}
pacman::p_load(tidyverse, readxl, haven, sjlabelled, BiocManager, kableExtra)
# create an object dataframe example `dfex` and assign to it the .sav file `sample.sav` that was introduced previously
dfex<-haven::read_sav("data/sample.sav")

# create an object movies metadata `dfmv` and assign to it the .xlsx file `movies.xlsx`
# note the different paths to these files
# note that we specify which sheet to read too; here only sheet 1 is imported
dfmv<-readxl::read_excel("data/movies.xlsx",1)

# next, we check if the source material was imported successfully by observing the first lines in the tables
head(dfex)
head(dfmv)

```

This is an example of how automatization can be implemented in the work flow. My list of movies include `r nrow(dfmv)` entries. The title of those movies are `r dfmv$Movie`. Is there a movie that I actually don't like on that list, well, the answer is that I dislike exactly `r dfmv %>% filter(Like %in% c("No","no","NO")) %>% nrow()` movies on that list.  
On average I liked all movies on a grade of `r mean(dfmv$Grade)` with the lowest rating of `r min(dfmv$Grade)` to the max rating of  `r max(dfmv$Grade)`. 

```{r, echo = F}
# complete cases
dfex_full <- dfex[complete.cases(dfex$age, dfex$gen, dfex$men_warm, dfex$men_comp),]

reg1 <- lm(men_warm ~ age + gen, data = dfex_full )
summary_reg1 <- summary(reg1)
coefficients <- summary_reg1$coefficients

#p_values <- coefficients[, "Pr(>|t|)"]
x <- "not significant"
y <- "significant"


```

The regression shows that age have an effect of `r coefficients[2,1]` with a p-value of `r coefficients[2,4]`. The estimator is `r ifelse(coefficients[2,4]< .05, y ,x)`.
Gender has an effect of `r coefficients[3,1]` with a p-value of `r coefficients[3,4]`. The estimator is `r ifelse(coefficients[2,4]< .05, y ,x)`.


*** 
## New Movie Tabels and Plots 

```{r}
dfmv %>% knitr::kable(caption="Simple table using knitr::kable()",format = "pipe")

# does some data manipulation to retrieve the required information
tmptbl<-dfmv %>% 
  filter(Actor %in% c("Keanu Reeves", "Alec Baldwin"))

# creates an empty table holder that is our summary table that we'd
# want to include in the final output document
extbl<-tibble(
  
  like=tmptbl[ tmptbl$Grade >= 8 |tmptbl$Grade <= 5 & tmptbl$Like %in% c("Yes","No"), ]$Like,
  name=tmptbl[ tmptbl$Grade >= 8 |tmptbl$Grade <= 5 & tmptbl$Like %in% c("Yes","No"), ]$Actor,
  movie=tmptbl[ tmptbl$Grade >= 8 |tmptbl$Grade <= 5 & tmptbl$Like %in% c("Yes","No"), ]$Movie,
  wiki=tmptbl[ tmptbl$Grade >= 8 |tmptbl$Grade <= 5 & tmptbl$Like %in% c("Yes","No"), ]$Wikilink
  )
```

```{r}
extbl %>% knitr::kable(caption="Movies graded 8 or more from liked and least like actors with Keanu Reeves and Alec Baldwin", format="pipe")
```

```{r}
# we remove all the labels using the package sjlabelled and mutate
# as factors columns gen and res
# this step makes it easier later on to work with plots
# note that it is not a necessary step in general but only for the sake of 
# simplicity here
dfex<-dfex %>% sjlabelled::remove_all_labels() %>% 
  mutate(gen=factor(gen),
         res=factor(res))

# subsamples 15 study participants at random
tmpdf1<-sample_n(dfex,15)
# subsample 60 study participants at random
tmpdf2<-sample_n(dfex,60)

```

```{r, warning=F, message=F}
haven::write_sav(tmpdf1,"data/tmpdf1.sav")
haven::write_sav(tmpdf2,"data/tmpdf2.sav")
```

```{r, warning = F, message=F}
# import dataset into one object and then subject this object to the ggplot code

# 1 - imports dataset into object tempdf
tempdf<-haven::read_sav("data/tmpdf1.sav") %>% sjlabelled::remove_all_labels() %>% 
  mutate(gen=factor(gen),
         res=factor(res))

# 2 - applies the ggplot to the dataset
ggplot(tempdf, aes(x=gen, y=wom_warm)) + 
  labs(x="Gender",
       y="Stereotype of warmth") +
  geom_boxplot() + 
  theme_light()



# 3 - for illustration purposes, repeat step 1 with each 
# of the three datasets (tmpdf1,tmpdf2 and dfex) 
# making sure they are assigned into the same object tempdf. 
# As long as the ggplot code is applied to a dataset with the same structure
# and variable labels the output will be updated automatically.
```

```{r, warning = F, message=F}
# import dataset into one object and then subject this object to the ggplot code

# 1 - imports dataset into object tempdf
tempdf<-haven::read_sav("data/tmpdf2.sav") %>% sjlabelled::remove_all_labels() %>% 
  mutate(gen=factor(gen),
         res=factor(res))

# 2 - applies the ggplot to the dataset
ggplot(tempdf, aes(x=gen, y=wom_warm)) + 
  labs(x="Gender",
       y="Stereotype of warmth") +
  geom_boxplot() + 
  theme_light()



# 3 - for illustration purposes, repeat step 1 with each 
# of the three datasets (tmpdf1,tmpdf2 and dfex) 
# making sure they are assigned into the same object tempdf. 
# As long as the ggplot code is applied to a dataset with the same structure
# and variable labels the output will be updated automatically.
```

```{r, warning = F, message=F}
# import dataset into one object and then subject this object to the ggplot code

# 1 - imports dataset into object tempdf
tempdf<-haven::read_sav("data/sample.sav") %>% sjlabelled::remove_all_labels() %>% 
  mutate(gen=factor(gen),
         res=factor(res))

# 2 - applies the ggplot to the dataset
ggplot(tempdf, aes(x=gen, y=wom_warm)) + 
  labs(x="Gender",
       y="Stereotype of warmth") +
  geom_boxplot() + 
  theme_light()



# 3 - for illustration purposes, repeat step 1 with each 
# of the three datasets (tmpdf1,tmpdf2 and dfex) 
# making sure they are assigned into the same object tempdf. 
# As long as the ggplot code is applied to a dataset with the same structure
# and variable labels the output will be updated automatically.
```
