---
title: "Parameterized Report"
author: "Charlotte Haussmann"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
params:
  agegroup:
    label: Agegroup
    value: "19-28"
    input: select
    choices:
    - "19-28"
    - "29-38"
    - "39-48"
    - "49-58"
    multiple: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F, warning=FALSE, message=FALSE}
# packages 
r <- getOption("repos")
r["CRAN"] <-"https://cloud.r-project.org/"
options(repos=r)

if (!require(pacman)) {
    install.packages("pacman")
    require(pacman)
}

pacman::p_load(tidyverse, ggplot2, haven, sjlabelled, dplyr, tidyr)
```

```{r, echo = F}
# read in data
df <- read_sav("data/sample.sav")
```

```{r, echo=F}
#preparation agegroup 
#["19-28", "29-38", "39-48", "49-58"] 

df$Agegroup <- df$age
df$Agegroup[df$age <= 28] <- "19-28"
df$Agegroup[df$age  >= 29 & df$age <= 38] <- "29-38"
df$Agegroup[df$age >= 39 & df$age <= 48] <- "39-48"
df$Agegroup[df$age >= 49] <- "49-58"


```

## Parameterized Report

```{r, echo = F}
ageg <- params$agegroup
ageg
```

```{r, echo = T}
tempdf <- sjlabelled::remove_all_labels(df) %>%
  filter(Agegroup %in% ageg)

agg_data <- tempdf %>%
  dplyr::count(age, gen)

ggplot(agg_data, aes(x = as.factor(ageg), y=n,  fill = factor(gen, levels = c(1,2), labels = c("female", "male")))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Agegroup", y = "Frequency", fill = "Gender", title = "Bar chart of age and gender") +
  theme_minimal()

```

First we look at some descriptive statistics. The age group `r ageg` has a sample size of `r nrow(tempdf)`.   


```{r, echo = T}
df %>% 
  sjlabelled::remove_all_labels() %>% 
  pivot_longer(contains("men_warm")) %>% 
  group_by(gen) %>% 
  summarise(mean=mean(value, na.rm = TRUE), 
            sd=sd(value, na.rm = TRUE),
            min=min(value, na.rm = TRUE),
            max=max(value, na.rm = TRUE))
```

```{r, echo = F}
df2 <- df %>% 
  sjlabelled::remove_all_labels() %>% 
  pivot_longer(contains("men_warm")) %>% 
  group_by(gen) %>% 
  summarise(mean=mean(value, na.rm = TRUE), 
            sd=sd(value, na.rm = TRUE),
            min=min(value, na.rm = TRUE),
            max=max(value, na.rm = TRUE))

df_clean <- df %>%
  sjlabelled::remove_all_labels() %>%
  pivot_longer(contains("men_warm"))

group1 <- df_clean %>% filter(gen == 1) %>% pull(value)
group2 <- df_clean %>% filter(gen == 2) %>% pull(value)

# Performing t-test
t_test_result <- t.test(group1, group2, na.rm = T)

p_value <- t_test_result$p.value
x <- "not significant"
y <- "significant"

```
In the table you can see that **men** evaluate the stereotypical evaluation of men in terms of **warmth** with a mean of `r df2[2,2]` and a standard deviation of `r df2[2,3]`. **Women** on the other hand evaluate the **warmth** of men with a mean of `r df2[1,2]` and a standard deviation of `r df2[1,3]`. The t-test shows that the differences in the means are `r ifelse(p_value< .05, y ,x)`.
